---
title: "ggplot2"
author: "Yujia Li"
date: "10/7/2021"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## data import same as Viz-1
1. *pull* _id_, _prcp_, _tmin_ and _tmax_ out of orginal data set; use _date_ to select certain ranges of observations.
2. *mutate* new variable _name_ from *recode*  _id_; and also redefine _tmin_ and _tmax_.  
```{r weather NYC}
weather_df = rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"), var = c("PRCP", "TMIN", "TMAX"), date_min = "2017-01-01", date_max = "2017-12-31") %>%
  mutate(name = recode(id, USW00094728 = "CentralPark_NY", USC00519397 = "Waikiki_HA",USS0023B17S = "Waterhole_WA"),tmin = tmin / 10,tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Labels
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scales for continuous and discrete
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")) +
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100, 300))
```

## Creating own colors 
## Themes for non-data modification
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE)

ggp_temp_plot + 
  theme_grey() + 
  theme(legend.position = "bottom")
```

## Merge different plots in ggplot
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## Patchwork
```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data Manipulation with factors
```{r}
weather_df %>%
  mutate(name = fct_reorder(name,tmax)) %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "obs", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = obs)) +
  geom_density(alpha = .5) + 
  facet_grid(.~ name)
```

## learning assessment
```{r}

```
