---
title: "ggplot1"
author: "Yujia Li"
date: "10/5/2021"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## data import
1. *pull* _id_, _prcp_, _tmin_ and _tmax_ out of orginal data set; use _date_ to select certain ranges of observations.
2. *mutate* new variable _name_ from *recode*  _id_; and also redefine _tmin_ and _tmax_.  
```{r weather NYC}
weather_df = rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"), var = c("PRCP", "TMIN", "TMAX"), date_min = "2017-01-01", date_max = "2017-12-31") %>%
  mutate(name = recode(id, USW00094728 = "CentralPark_NY", USC00519397 = "Waikiki_HA",USS0023B17S = "Waterhole_WA"),tmin = tmin / 10,tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Plots
How to *color*, add *smooth curve* and change the *transparency* in ggplot
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

## The difference in adding **color** to ggplot and geom 
## facet_grid argument with *.~variable* 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## The precipitation in NYC
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Compare histogram and density 
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

## box plot
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

## ridge plot
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

## Learning assessment
use _weather_df_, only with Central Park data;
converts temp to Fahrenheit;
makes a scatter plot of min vs. max temp
overlays a linear regression 
```{r}
weather_df_LA = filter(weather_df, name == "CentralPark_NY") %>%
  mutate(tmax = 9/5*tmax + 32, tmin = 9/5*tmin + 32) %>%
  ggplot(aes(x = tmin, y = tmax)) + geom_point(alpha = .3) + geom_smooth(se = FALSE, method = "lm")

plot(weather_df_LA)
```

## Saving the plot 
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)

knitr::opts_chunk$set(fig.width = 6, fig.asp = .6,out.width = "90%")
```