---
title: "Visulization_3 Explanatory data analysis"
author: "Yujia Li"
date: "10/12/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(fig.width = 6,fig.asp =  .6, out.width = "90%")

theme_set(theme_minimal(),theme(legend.position = "bottom"))

options(ggplot2.continuous.color = "viridis", ggplot2continuous.fill = "viridis")

scale_color_discrete = scale_colour_viridis_d
scale_color_discrete = scale_fill_viridis_d
```

## data import same as Viz-1 and Viz_2
1. *pull* _id_, _prcp_, _tmin_ and _tmax_ out of orginal data set; use _date_ to select certain ranges of observations.
2. *mutate* new variable _name_ from *recode*  _id_; and also redefine _tmin_ and _tmax_.  
*adding _lubridate_ package in the _mutate_ function
```{r weather NYC}
weather_df = rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"), var = c("PRCP", "TMIN", "TMAX"), date_min = "2017-01-01", date_max = "2017-12-31") %>%
  mutate(name = recode(id, USW00094728 = "CentralPark_NY", USC00519397 = "Waikiki_HA",USS0023B17S = "Waterhole_WA"),tmin = tmin / 10,tmax = tmax / 10, month = lubridate::floor_date(date,unit = "month")) %>%
  select(name, id, everything())
```

## _group by_ or _ungroup_
Now, you can check the weather_df and see Groups:   name, month [36]
```{r}
weather_df %>% 
  group_by(name,month)
```

## Counting things 
Combining group_by and summarize, n() or in base R fashion using count(). Both produce a tibble.
*Never use base R table
```{r}
weather_df %>% 
  group_by(name,month) %>%
  summarize(n_obs = n())

weather_df %>% 
  count(name, month, name = "n_obs")
```

## Count distinct things
n_distinct looks into how many days in the month
```{r}
weather_df %>% 
  group_by(month) %>%
  summarize(n_obs = n(),n_days = n_distinct(date))
```

## 2x2 table
Janitor::tabyl
```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

## general summary
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

## Summarize for ggplot
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

## Summarize for table
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

## grouped mutate
Adding a new column for *mean of tmax variable*. Then you can center tmax by using the new variable *mean of tamx* to center temp around 0.
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

## Window function 
ranking with _mutate_ or _filter_ function 
```{r}
weather_df %>%
  group_by(name,month) %>%
  mutate(temp_rank = min_rank(tmax))

weather_df %>%
  group_by(name,month) %>%
  filter(min_rank(tmax) < 2)
```

## lag
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

```{r}
```
